---
title: "Assignment 3"
author: "Raju Adhikari and Mario Alonso Rodriguez"
date: "11/11/2016"
output: 'pdf_document'
---

```{r options, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, message=FALSE)
```

## Merging databases

This is the third pair assignment of Introduction to Collaborative Social Science Data Analysis. Here, we 
gathered web based data from the World Bank API, as well as from the Emissions Database for Global Atmospheric Research (EDGAR) from the European Comission website. With both datasets we aim to answer our research question: Does income inequality affect the environment negatively? To operationalize this, we will use CO2 emissions per capita to measure effects on the environment (dependent variable) and the GINI coefficient for inequality (independent variable). We merged the two data sets, and conducted basic descriptive and inferential statistics to address this question. 

```{r databases, warning=FALSE, error=FALSE, message=FALSE, echo=FALSE, results='hide'}

####First of all, we set the working directory.

setwd('/Users/mariorodriguez/Desktop/Assignment3')

#### We load the packages we need for the handling of the data.
library(WDI)
library(countrycode)
library(rio)
library(stargazer)
library(ggplot2)

##### We access the World Development Indicators from the World Bank API direcly from R and download the GINI coefficient indicator, also GDP per capita in case we need it as a control variable. We use the data from 2010 since it is the last one with the GINI coefficient for the United States.

GINI <- WDI(country = 'all', start = '2010', end = '2010', indicator = c('SI.POV.GINI', 'NY.GDP.PCAP.CD'), extra = TRUE)

##### We download the dataset of CO2 emissions per capita from the EDGAR database, unfortunately it is an xls file so we need to use the rio package.

CO2 <- import('http://edgar.jrc.ec.europa.eu/news_docs/CO2_per_capita_1970-2014_dataset_of_CO2_report_2015.xls')

##### We check the CO2 dataframe that we just downloaded and see that we don't need the first 13 rows so we take them out. Also, we only need the last column with the information of the CO2 emissions of 2013. We label that column since it does not have a name from the import process.

CO2 <- CO2[-c(1:13), ]
names(CO2)[42] <- "CO2010"

###### We also rename the first column with the countries and create a new variable with the iso2c codes of the countries so it's easier to merge the data. Also we get rid from all the columns with information of the previous years. 

names(CO2)[names(CO2)=="Global per capita CO2 emissions from fossil fuel use and cement production 1970-2014"] <- "country"

CO2$iso2c <- countrycode(CO2$country, 'country.name', 'iso2c', warn = TRUE)

CO2 <- CO2[, c('iso2c', 'CO2010')]

##### We take a look at the data that we downloaded from the World Bank and change the name of the variables that we are going to use.

View(GINI)
names(GINI)[names(GINI)=="SI.POV.GINI"] <- "GiniCoeff"
names(GINI)[names(GINI)=="NY.GDP.PCAP.CD"] <- "GDPperCapita"

##### We check the variable types and see that the iso2c is factor so we change it to character.

GINI$iso2c <- as.character(GINI$iso2c)

GINI <- GINI[, c('iso2c', 'country', 'GiniCoeff', 'GDPperCapita', 'income')]

#### We merge the datasets using iso2c as the base and check that the database has all the variables.

Combined <- merge(GINI, CO2, by = c('iso2c'))

View(Combined)


#### We remove the rows with NAs given that they will not be used on our inferential analysis and are left with 70 observations.

Combined <- Combined[complete.cases(Combined),]

```

## Descriptive statistics

A series of basic statistics show that the mean Gini is 36.81. The minimum value is 24.82 for the Ukraine, which means that it was the less unequal country in terms of income in 2010. The highest inequality was in Zambia with 55.62. The mean value of GDP per capita of our set of countries in 2012 was 19,222.2 USD dollars. The lowest value for this variable is 341.9 of Ethiopia and the highest is 103267.3 of Luxembourg. Finally, the CO2 emissions per capita have a mean of 4.873 metric tons in 2012. The lowest value in 2010 was 0.056 of Tuvalu, and the highest is 21.941 from Luxembourg.

```{r descriptive, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
#### We create a table to get all the descriptive statistics of the newly created dataset.

stargazer(Combined, type = 'text')


```

## Inferential statistics

A plot shows that there might not be a linear relationship between inequality and CO2 emissions. We run a first regression model with only one explanatory variable and see that it is statistically significant and that actually, more equal countries (with lower GINI coefficients) are the ones that emit more CO2. A one unit increase in the Gini coefficients means a decrease in the emission of about 0.3 metric tons of CO2 anually. This could be to the fact that countries with less inequality are more developed. On the second model, we include a control variable of GDP per Capita to see if the relationship between our independent and dependent variable is not caused by this third variable. The second model shows that GDP per capita and the Gini coeficcient are statistically significant. Nonetheless, our inequality variable is still negative, which means that countries with higher inequality actually have less CO2 emissions, even when controling for income. 

```{r inferential, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

ggplot(data = Combined, aes(x = CO2010,
        y = GiniCoeff)) + geom_point(aes(colour = income)) + theme_bw()

m1 <- lm(CO2010 ~ GiniCoeff, Combined)

m2 <- lm(CO2010 ~ GiniCoeff + GDPperCapita, Combined)

stargazer(m1, m2, type = 'text')
```

