---
title: "Assignment 3"
author: "Raju Adhikari and Mario Alonso Rodriguez"
date: "November 11, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Merging databases

This is the third pair assignment of Introduction to Collaborative Social Science Data Analysis. Here, we 
gathered web based data from the World Bank API, as well as from the World Health Organization website. We merged the data sets, and conducted basic descriptive and inferential statistics to address the effect of inequality on environment.

```{r cars}
install.packages('WDI')
install.packages('countrycode')
library(WDI)
library(countrycode)
library(plyr)
GINI <- WDI(country = 'all', start = '2012', end = '2012', indicator = 'SI.POV.GINI', extra = TRUE)
DEATHS <- read.csv('http://apps.who.int/gho/athena/data/xmart.csv?target=GHO/TOTENV_1,TOTENV_2,TOTENV_5&profile=crosstable&filter=COUNTRY:*;REGION:*;ENVCAUSE:ENVCAUSE000;ENVCAUSE:ENVCAUSE001;ENVCAUSE:ENVCAUSE060;ENVCAUSE:ENVCAUSE151&x-sideaxis=COUNTRY;YEAR&x-topaxis=GHO;ENVCAUSE')
DEATHS <- DEATHS[-c(1), ]
print(class(DEATHS$X))
DEATHS$X <- as.character(DEATHS$X)
names(DEATHS)[names(DEATHS)=="X"] <- "country"
names(DEATHS)[names(DEATHS)=="X.1"] <- "year"
names(DEATHS)[names(DEATHS)=="Deaths.attributable.to.the.environment...."] <- "Percentageofdeaths"
DEATHS$iso3c <- countrycode(DEATHS$country, origin = 'country.name', destination = 'iso3c', warn = TRUE)
DEATHS <- DEATHS[, c('iso3c', 'year', 'Percentageofdeaths')]
names(GINI)

GINI <- GINI[, c('iso3c', 'country', 'SI.POV.GINI')]

Combined <- merge(GINI, DEATHS, by = c('iso3c'))


```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
